---
title: "BIS15L_AI_class"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Today, we're exploring how AI can aid in writing, comprehending and troubleshooting our R code. AI is a very valuable tool, but it's important not to rely on it entirely. Building a strong R foundation is still important. These AI tools are not perfect and can be inaccurate solutions. In addition these tools can't read our minds, so understanding R helps us generate better prompts to get better responses. In short, AI should be used as a tool, not a crutch. Additionally, do not forget about other useful resources like Stack Overflow/Exchange, online tutorials, and GitHub issues pages for learning and troubleshooting. Let's combine AI with traditional methods to master R effectively.

This overview isn't exhaustive of all possible AI tools for R coding, but I believe it is a good start and covers everything you need.


**Make sure to enable Github Copilot in RStudio.**
Go to Tools -> Global Options -> Copilot -> Enable Git

## Learning Goals: 
*At the end of this exercise, you will be able to:*  
1. Explain what AI is and what AI tools can be used for R.
2. Enable Github Copilot in RStudio.
3. Use Copilot/ChatGPT to write code in RStudio.
4. Use Copilot/ChatGPT for troubleshooting code in RStudio.


## Exploration Questions:
Lets get some get a feel for what copilot (or ChatGPT) does, use Github Copilot to autocomplete the questions below. You can alternatively paste the questions in ChatGPT or Microsoft Copilot.
1. What is AI?
  AI is the simulation of human intelligence processes by machines, especially computer systems. These processes include learning, reasoning, and self-correction. Specific applications of AI include expert systems, natural language processing (NLP), speech recognition and machine vision. 
  
2. What is ChatGPT?
ChatGPT is a language model developed by OpenAI. It is a large-scale transformer-based language model that can generate human-like text. It is trained on diverse internet text data and can generate human-like responses to prompts.


3. What is Github Copilot?
Github Copilot is an AI pair programmer that helps you write code faster. It is powered by OpenAI's GPT-3 and is available as a plugin for Visual Studio Code and RStudio. It can help you write code, suggest code completions, and help you find solutions to coding problems.

4. [Try it yourself: Insert your own question here]
How did World War 1 start? 
World war 1 started in 1914 when Archduke Franz Ferdinand of Austria was assassinated by a Serbian nationalist. This event led to a series of events that eventually led to the outbreak of the war.

## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(dplyr)
library(shiny)
library(shinydashboard)
```

## Load the Data: 
```{r}
#load the Netflix Data in your Data folder, save it under the object name netflix

netflix <- read_csv("Data/netflix_titles.csv")


```

*Did copilot help you write the code above? Could anything about the prompt be improved to help Copilot understand what you want to do more accurately?*


## Lets Practice:
Explore the elements in your data: include a comment within your code chunk to prompt Copilot to use the function of your choice
```{r}
#explore the data 
head(netflix)
glimpse(netflix)

```
*Did copilot use your preferred method to get the summary of the data? If not, how could you improve the prompt to get the desired output?*


Sort out the movies from the netflix into an new object called netflix_movies
```{r} 
#sort out the movies from the netflix data
netflix_movies <- netflix %>% filter(type == "Movie")

```


Which director has the most movies on netflix?
```{r}
#find director with the most movies on netflix
netflix_movies %>% 
  filter(director != "NA") %>%
  count(director, sort = TRUE)



```
*Do you see anything wrong with the table above, look at row 1. Write a prompt below to fix this.*

Which director has the most movies on netflix? Filter out NA's
```{r}
#Which director has the most movies on netflix? Filter out NA's 
netflix_movies %>% 
  filter(!is.na(director)) %>%
  count(director, sort = TRUE)
```

Which director had the most movies by release year and remove the NA's?
```{r}
#Which director had the most movies by release year and remove the NA's?
netflix_movies %>% 
  filter(!is.na(director)) %>%
  count(director, release_year, sort = TRUE)

```
*See anything else interesting about this table, look at the first director on the table. Could this be a problem and how can we edit the dataset to fix this? Try writing a prompt.*


Which director had the most movies by release year and remove the NA's? Separate commas in the director column

```{r}
#Which director had the most movies by release year and remove the NA's? Seperate commas in the director column 
netflix_movies <- netflix_movies %>%
  separate_rows(director, sep = ", ") %>%
  filter(!is.na(director)) %>%
  count(director, release_year, sort = TRUE)

```


Make a bar plot of the top directors with the most movies on netflix. 
```{r}
#make a bar plot of the top directors with the most movies on netflix, fill by director. Remove the legend flip axis and make the title "Top 10 Directors"

netflix_movies %>% 
  filter(!is.na(director)) %>%
  count(director, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x = fct_reorder(director, n), y = n, fill = director)) +
  geom_col() +
  coord_flip() +
  labs(title = "Top 10 Directors with the Most Movies on Netflix", 
       x= "Directors",
       y= "Number of Movies") +
  theme(legend.position = "none")






```
*Do you see any issues with this plot? How could we improve the prompt to get a better boxplot? Try writing a prompt.*
```{r}
#make variable year a factor 
netflix_movies$release_year <- as.factor(netflix_movies$release_year)
```

[Insert your prompt here]
```{r}
netflix_movies %>% 
  filter(filter == 2020) %>%
  filter(!is.na(director)) %>%
  count(director, sort = TRUE) %>%
  top_n(10) %>%
  ggplot(aes(x = fct_reorder(director, n), y = n, fill = director)) +
  geom_col() +
  coord_flip() +
  labs(title = "Top 10 Directors with the Most Movies on Netflix", 
       x= "Directors",
       y= "Number of Movies") +
  theme(legend.position = "none")

```

*How about for just 2020*

```{r}
#make a bar plot of the top directors with the most movies on netflix in 2020, fill by director. Remove the legend flip axis and make the title "Top 10 Directors in 2020" 

netflix_movies %>% 
  filter(!is.na(director)) %>%
  filter(release_year == 2020) %>%
  count(director, sort = TRUE)%>%
  top_n(10, director) %>%
  ggplot(aes(x = fct_reorder(director, n), y = n, fill = director)) +
  geom_col() +
  coord_flip() +
  labs(title = "Top 10 Directors with the Most Movies on Netflix in 2020", 
       x= "Directors",
       y= "Number of Movies") +
  theme(legend.position = "none")


```
[Insert your prompt here]
```{r}

```


*Take a few minutes to make your own plot(s) using prompts, edit prompts until you get a plot that you like.*






## Lets make the top 10 directors barplot into a Shiny App: use copilot or Chatgpt
```{r}
#make the top 10 directors barplot into an interactive Shiny App for all years. end sessions when user closes app
# fix code to make it run faster and more efficiently
 

ui <- fluidPage(
  titlePanel("Top 10 Directors with the Most Movies on Netflix"),
  plotOutput("top_directors")
)

server <- function(input, output) {
  output$top_directors <- renderPlot({
    netflix_movies %>% 
      filter(!is.na(director)) %>%
      count(director, sort = TRUE) %>%
      top_n(10) %>%
      ggplot(aes(x = fct_reorder(director, n), y = n, fill = director)) +
      geom_col() +
      coord_flip() +
      labs(title = "Top 10 Directors with the Most Movies on Netflix", 
           x= "Directors",
           y= "Number of Movies") +
      theme(legend.position = "none")
  })
}

shinyApp(ui = ui, server = server)


```



*Now use ChatGPT, Copilot or previous notes to fix the code*





## That's it! Let's take a break and then move on to working on group projects.
