---
title: "Lab 7 Homework"
author: "Your Name Here"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the libraries
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(skimr)
library(dplyr)
```

For this assignment we are going to work with a large data set from the [United Nations Food and Agriculture Organization](http://www.fao.org/about/en/) on world fisheries. These data are pretty wild, so we need to do some cleaning. First, load the data.  

Load the data `FAO_1950to2012_111914.csv` as a new object titled `fisheries`.
```{r}

fisheries <- readr::read_csv(file = "data/FAO_1950to2012_111914.csv")
```

1. Do an exploratory analysis of the data (your choice). What are the names of the variables, what are the dimensions, are there any NA's, what are the classes of the variables?  
```{r}
dim(fisheries)
names(fisheries)
anyNA(fisheries)
```

2. Use `janitor` to rename the columns and make them easier to use. As part of this cleaning step, change `country`, `isscaap_group_number`, `asfis_species_number`, and `fao_major_fishing_area` to data class factor. 
```{r}
new_fisheries <- clean_names(fisheries)
new_fisheries$country <- as.factor(new_fisheries$country)
new_fisheries$isscaap_group_number  <- as.factor(new_fisheries$isscaap_group_number )
new_fisheries$asfis_species_number  <- as.factor(new_fisheries$asfis_species_number )
new_fisheries$fao_major_fishing_area  <- as.factor(new_fisheries$fao_major_fishing_area )
new_fisheries$year
```

We need to deal with the years because they are being treated as characters and start with an X. We also have the problem that the column names that are years actually represent data. We haven't discussed tidy data yet, so here is some help. You should run this ugly chunk to transform the data for the rest of the homework. It will only work if you have used janitor to rename the variables in question 2!  
#this cleans data, remove the hashtags
```{r}
 fisheries_tidy <- new_fisheries %>% 
  pivot_longer(-c(country,common_name,isscaap_group_number,isscaap_taxonomic_group,asfis_species_number,asfis_species_name,fao_major_fishing_area,measure),
               names_to = "year",
               values_to = "catch",
               values_drop_na = TRUE) %>% 
  mutate(year= as.numeric(str_replace(year, 'x', ''))) %>% 
  mutate(catch= str_replace(catch, c(' F'), replacement = '')) %>% 
  mutate(catch= str_replace(catch, c('...'), replacement = '')) %>% 
  mutate(catch= str_replace(catch, c('-'), replacement = '')) %>% 
  mutate(catch= str_replace(catch, c('0 0'), replacement = ''))

fisheries_tidy$catch <- as.numeric(fisheries_tidy$catch)
```

3. How many countries are represented in the data? Provide a count and list their names.
```{r}
fisheries_tidy %>%
  summarise(number_countries=n_distinct(country))
```

4. Refocus the data only to include country, isscaap_taxonomic_group, asfis_species_name, asfis_species_number, year, catch.
```{r}
specific_fish <- fisheries_tidy %>%
  select(country, isscaap_taxonomic_group, asfis_species_name, asfis_species_number, year, catch)

```

5. Based on the asfis_species_number, how many distinct fish species were caught as part of these data?
```{r}
fisheries_tidy %>%
  summarise(number_species = n_distinct(asfis_species_number))
```
    1551 distict fish species were caught in this dataset
    
6. Which country had the largest overall catch in the year 2000?
```{r}
fisheries_tidy %>%
  select(country, catch, year) %>%
  filter(year == 2000) %>%
  arrange(desc(catch))
```
    China had the largest catch in 2000, with a catch of 9068 tons
    
7. Which country caught the most sardines (_Sardina pilchardus_) between the years 1990-2000?
```{r}
fisheries_tidy %>%
  group_by(country) %>%
  select(country, asfis_species_name, catch, year) %>%
  filter(year >= 1990 & year <= 2000) %>%
  filter(asfis_species_name == "Sardina pilchardus") %>%
  summarise(total_catch= sum(catch))%>%
  arrange(desc(total_catch))
```
    Morocco caught the most sardines between the years 1990-2000, with a total catch of 7470 tons caught.

8. Which five countries caught the most cephalopods between 2008-2012?
```{r}
fisheries_tidy %>% 
  group_by(country) %>%
  select(country, asfis_species_name, catch, year) %>%
  filter(asfis_species_name == "Cephalopoda") %>%
  filter(year >= 2008 & year <= 2012) %>%
  summarise(total_ceph = sum(catch)) %>%
  arrange(desc(total_ceph))
```
    India (570), China (257), Algeria (162), France (101), and TimoeLeste (76) caught the most cephalapods between 2008 and 2012

9. Which species had the highest catch total between 2008-2012? (hint: Osteichthyes is not a species)
```{r}
fisheries_tidy %>%
  group_by(asfis_species_number) %>%
  select(year, asfis_species_number, catch, common_name) %>%
  filter(year >= 2008 & year <= 2012) %>%
  summarise(most_caught = sum(catch)) %>%
  left_join(fisheries_tidy  %>%
              filter(year >= 2008 & year <= 2012) %>%
              distinct(asfis_species_number, common_name),
            by = "asfis_species_number" ) %>%
  arrange(desc(most_caught))
fisheries_tidy %>%
  subset(asfis_species_number == "1480401601")

  
```
    Alaska Pollock, Peruvian anchovy, Pacific saury, Indian oil sardine, and California plichard were the most caught species between 2008 and 2012.
 
 
10. Use the data to do at least one analysis of your choice.
    
    I will be analzying what the country with the highest total catch amount is
    
```{r}
fisheries_tidy %>%
  group_by(country) %>%
  select(country, year, catch) %>%
  summarise(total_catch_country= sum(catch, na.rm =T)) %>%
  arrange(desc(total_catch_country))
  
```
    Japan has the highest total catch with an all time total of 814,843 tons of total catch.


## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   